\chapter{Assignment 9}
\section{Lecture 43.00: Stratified Sampling}
Suppose that you're interested in the population as a whole,
but at the same time you're interested in some sub-population of the population.
For example suppose you're interested in University of Waterloo students,
but we're also interested in Math faculty students.
You would use a \emph{stratified} sample.
In this case, you might perform SRS over a sub-population and then
combine that together to get the entire population so what would that notationally
look like?

Suppose we have a frame $ U $, and we divide $ U $ into sub-frames
$ U_1,U_2,\ldots,U_H $ where
\begin{enumerate}[(1)]
    \item $ U_1\cup U_2\cup U_3\cup \cdots \cup U_H=U $.
    \item For any $ i\ne j $, $ U_i\cap U_j=\emptyset $.
    \item Define $ \abs{U_h}=N_h $ and $ \abs{U}=N $.
    \item $ N_1+N_2+\cdots+N_h=N $.
\end{enumerate}
\subsubsection*{Parameter}
\begin{align*}
    \mu
     & =\frac{N_1 \mu_1+N_2 \mu_2+\cdots+N_H \mu_H}{N}                     \\
     & =\frac{N_1}{N} \mu_1+\frac{N_2}{N} \mu_2+\cdots+\frac{N_h}{N} \mu_H \\
     & =w_1 \mu_1+\cdots+w_H \mu_H                                         \\
     & =\sum_{i=1}^{H} w_i \mu_i
\end{align*}
\subsubsection*{Estimate}
We would use SRS to estimate the strata's average.
\[ \hat{\mu}=\sum_{i=1}^{H} w_i\hat{\mu}_i \]
When we get the strata's average,
and we multiply by the weight we add over the strata's to
get the population average.
\subsubsection*{Estimator}
\[ \tilde{\mu}=\sum_{i=1}^{H} w_i\tilde{\mu}_i \]
where $ \tilde{\mu}_i $ is the SRS random variable, hence it is unbiased
and normally distributed. Therefore, the weighted average of the estimator
will be normally distributed. Now, we need to find the expectation and variance
of $ \tilde{\mu} $.
\subsubsection*{Expectation of $ \tilde{\mu} $}
\[ \E*{\tilde{\mu}}=\E*{\sum_{i=1}^{H} w_i\tilde{\mu}_i}=\sum_{i=1}^{H} w_i\E{\tilde{\mu}_i}=
    \sum_{i=1}^{H} w_i \mu_i\quad\text{since SRS is unbiased.} \]
\subsubsection*{Variance of $ \tilde{\mu} $}
At this point in the course the formulas become ugly.
\[ \Var*{\tilde{\mu}}=\Var*{\sum_{i=1}^{H} w_i\tilde{\mu}_i}=\sum_{i=1}^{H} w_i^2\Var*{\tilde{\mu}_i} \]
where $ \tilde{\mu}_i\indep \tilde{\mu}_j $ since no unit is in both groups. Therefore,
\[ \Var*{\tilde{\mu}}=\sum_{i=1}^{H}w_i^2 \frac{\sigma_i^2}{n_i} \biggl(1-\frac{n_i}{N_i} \biggr) \]
\subsubsection*{Confidence Interval for $ \mu $}
\[ \hat{\mu}\pm c\sqrt{\sum_{i=1}^{H}w_i^2 \frac{\sigma_i^2}{n_i} \biggl(1-\frac{n_i}{N_i} \biggr)}\quad\text{where $ c \sim \N{0,1} $.} \]
\subsection*{Proportion}
For a proportion, we want to estimate $ \pi $.
\subsubsection*{Parameter}
\[ \pi=\sum_{i=1}^{H} w_i \pi_i \]
\subsubsection*{Estimate}
\[ \hat{\pi}=\sum_{i=1}^{H} w_i\hat{\pi}_i\quad\text{which uses SRS.} \]
\subsubsection*{Estimator}
\[ \tilde{\pi}=\sum_{i=1}^{H} w_i\tilde{\pi}_i \]
\subsubsection*{Confidence Interval for $ \pi $}
\[ \hat{\pi}\pm c\sqrt{\sum_{i=1}^{H}w_i^2 \frac{\sigma_i^2}{n_i} \biggl(1-\frac{n_i}{N_i} \biggr)}\quad\text{where $ c \sim \N{0,1} $.} \]
You'll remember that we have worked out before that $ \sigma_i^2=\pi_i(1-\pi_i) $.
