\chapter{Introduction}
\section{Preamble}
\begin{itemize}
    \item The subject matter of \emph{computational} statistics
          is that of \emph{Statistics} itself, but developed via computation
          rather than only \emph{through mathematics}.

          {\color{blue} Essentially, we are rebuilding our understanding of statistics
          through a computational perspective to be more general.}

    \item The goal of the course is to present essential statistical concepts.
          \begin{itemize}
              \item Simulation is used to illustrate the concepts and to provide understanding.
              \item Mathematical development provides an alternative presentation of the same
                    ideas, when that is possible, and is used to develop a tool or get insight into
                    a concept.
          \end{itemize}
    \item Because simulation is the primary means to develop this understanding.
          \begin{itemize}
              \item The statistics/estimators/tests/etc.\ used should be
                    of sufficient complexity that a complete mathematical treatment would be
                    beyond the level of this course.
          \end{itemize}
    \item Because the statistics/estimators/tests/etc.\ can be complex, several
          numerical methods are introduced. For example:
          Gradient descent, Newton's method, iteratively re-weighted squares, etc.
\end{itemize}
\subsection{Programming in R}
\begin{itemize}
    \item Students are expected to program in R\@; other languages will not
          be accepted.
    \item The ideas is to convey programming concepts alongside the statistical concepts.
          The purpose of the programming, like mathematics, is to illustrate the statistical
          concepts.
          \begin{itemize}
              \item Therefore, we will take advantage of the functional programming language
                    R to write some general purpose code.
          \end{itemize}
    \item The clarity and simplicity of the code is primary since it is being used
          to convey the statistical concepts; that is, the code is not
          production-level code, but instead teaching-level code.
    \item Wh
\end{itemize}
