\makeheading{Lecture 12}
\textbf{Example - Error Correcting Code}

Suppose you send a $ 4 $-bit message over a noisy connection.
Each bit is independently flipped with a probability $ p=0.1 $. Find
$ P(\text{message is received correctly}) $.

\textbf{Solution.}

Let $ X= $ \# of bits that get flipped.
\[ X \sim \bin(4,0.1) \]
\[ P(X=0)=\binom{4}{0}0.1^0(1-0.1)^{4-0}=0.6561 \]

\textbf{Example cont.}

Now suppose we add $ 3 $ parity bits that allow the receiver
to detect and correct up to $ 1 $ error in the message. Find
$ P(\text{message is received correctly}) $.

\textbf{Solution.}

Let $ Y= $ \# of bits flipped.
\[ Y \sim Bin(7,0.1) \]
\[P(Y=0)+P(Y=1)= \binom{7}{0}0.1^0 0.9^7 + \binom{7}{1}0.1^1 0.9^6\approx 0.8503 \]

\textbf{Example (ION)}

Suppose there are $ 100 $ people, $ 10 $ people with no bus pass,
and $ 20 $ people are selected at random without replacement.
Find the probability that there are two people with no bus pass.

\textbf{Solution.}

Let $ X= $ \# people found with no bus pass.
\[ X \sim \hyp(100,10,20) \]

\[ P(X=2)=\frac{\binom{10}{2}\binom{90}{18}}{\binom{100}{20}}\approx 0.3182 \]

\textbf{Hypergeometric Approximation with Binomial Distribution}

If we did the selection with repetition, the Hypergeometric distribution
would become a Binomial distribution.

If $ N $ is extremely large compared to $ n $, then it doesn't make much difference
if the sampling is with our without.

In that case, a $ \hyp(N,r,n) $ case can be approximated by $ \bin(n,\sfrac{r}{N}) $

\textbf{Example}

Suppose we have $ 100 $ cards, with $ 70 $ money cards and $ 30 $ non-money cards.
Select $ 4 $ cards without replacement.
Find a suitable approximation for this Hypergeometric distribution assuming
$ Z= $ \# of money cards selected.

\textbf{Solution.}

Let $ Z= $ \# of money cards selected.
\[ Z \sim \hyp(100,70,5) \]
\[ P(Z=4)=\frac{\binom{70}{4}\binom{30}{1}}{\binom{100}{5}}\approx 0.3654 \]

With a Binomial approximation:
\[ Z \sim \bin(5,\sfrac{70}{100}=0.7) \]
\[ P(Z=4)=\binom{5}{4}0.7^4(1-0.7)^{5-4}= 0.36015 \]
which is clearly a very good approximation.
