\makeheading{Lecture 13}

\section{Negative Binomial Distribution}
\begin{defbox}
    \subsection{Definition (Negative Binomial Distribution)}
    Suppose an experiment follows the Bernoulli trials, and continue
    doing this experiment until $ k $ success are obtained. Let $ X $
    be the number of failures obtained before the $ k^{\text{th}} $
    success. Then $ X $ has a \emph{Negative Binomial distribution}
    and we write
    \[ X \sim \nb(k,p) \]
\end{defbox}

\textbf{Find the probability function, $ f(x)$}

There will be a total of $ (x+k) $ trials ($ x $ F's $ k $ S's)
and the last trial will be a success. In the first $ (x+k-1) $ trials
we must have $ x $ failures and $ (k-1) $ successes, where order does not matter
(the last trial we know it must be a success, that's why we are only looking at
before the $ k^{\text{th}} $ success!)
Hence, we have a combination which is $ \binom{x+k-1}{k-1} $. Each order
will have a probability $ p^k(1-p)^x $. Hence
\[ f(x)=\binom{x+k-1}{k-1}p^k(1-p)^x \]
for $ x\in\rinterval{0}{\infty} $.

In a picture:
\[\overbrace{
        \underbrace{
            \text{\_ \_ \_ $\ldots$ \_}\mid}
        _{x \text{ F's},\,(k-1) \text{ S's}}}^{x+(k-1)\text{ Trials}}
    \underbrace{ \text{ S } }_{ k^{\text{th}} \text{ S}} \]

\textbf{Example}

How many tails until we get the $10^{\text{th}}$ head on a fair coin?

\textbf{Solution.}

Let $ X= $ \# of tails before the $ 10^{\text{th}} $ head.
\[ X\sim\nb\left(10,\sfrac{1}{2}\right) \]

\textbf{Example}

If courses were independent with probability $ p $ of passing
and you need $ 40 $ courses, then the number of failed courses
would be $ \nb(40,p) $.

\textbf{Example}

Suppose a start-up is looking for $ 5 $ investors. They ask
investors repeatedly where each independently has a $ 20\% $ chance
of saying yes. Let $ X= $ total \# of investors that they ask and
note that $ X $ does not follow a negative binomial distribution.
Find $ f(x) $ and $ f(10) $.

\textbf{Solution.}

Let $ Y= $ \# who say no before $ 5 $ say yes. We know $ X=Y+5 $.
\[ Y\sim\nb(5,0.2) \]
\begin{align*}
    f(x) & =P(X=x)                                   \\
         & =P(Y+5=x)                                 \\
         & =P(Y=x-5)                                 \\
         & =\binom{(x-5)+5-1}{5-1}(0.2)^5(0.8)^{x-5} \\
         & =\binom{x-1}{4}(0.2)^5(0.8)^{x-5}
\end{align*}
for $ x\in\rinterval{5}{\infty} $.
\[ f(10)=\binom{9}{4}(0.2)^5(0.8)^5\approx 0.0132\]
Note that we have $ \binom{9}{4} $ and not $ \binom{10}{5} $ because
the $10^{\text{th}}$ investor must have said yes.
