\makeheading{2019-10-08}
\subsection{Convergence of Simplex Algorithm}
In each iteration, we choose $k\in N$ such that $c_k>0$. Then, we compute
$t=\min\{\sfrac{b_i}{A_{ik}}\mid A_{ik}>0\}$. Then, throughout the
rest of the Simplex iterations, we never see the same basis again.
There are at most $\binom{n}{m}$ bases of A. Therefore, if $t>0$ in each iteration,
the Simplex Algorithm will terminate in at most $\binom{n}{m}$ iterations. The only
way the algorithm will not terminate is when $t=0$ for all iterations
(after some \# of iterations). If our choices for $k$ and $l$ are deterministic and
consistent in this case if we repeat a basis we call it a \emph{cycle}.

\begin{thmbox}
    \begin{theorem}
        The Simplex Algorithm starting from a basic feasible solution with
        Bland's Rule terminates.
    \end{theorem}
\end{thmbox}

\subsection{Implementation of the Simplex Algorithm in ``Big Data''}
In a given iteration of the Simplex Algorithm, what information do we need to
execute the algorithm?


We have the original data ($A$, $\bm{b}$, $\bm{c}$) and we have the
current $B$, $\bm{\bar{x}}$, $\bm{\bar{v}}$.

Pick any $k\in N$ such that $\bar{c_k}\geqslant  0$. $\bar{c_k}=c_k-\bm{\bar{y}}^\top  A\bm{x}$
(where $\bm{\bar{y}}^\top =c_B A_B^{-1}$).

Then to compute $t$, we need
$t=\min\{\sfrac{b_i}{A_{ik}}\mid A_{ik}>0\}$.

So, we need
$\bar{A_k}$: $\bar{A_k}=A_B^{-1}A_k$ and note that $\bm{\bar{x_N}}=\bm{0}$,
$\bm{\bar{x_B}}=\bm{\bar{b}}$


We solve linear systems $A_B^\top \bm{y}=\bm{c_B}$ and $A_B\bm{d_B}=A_k$.

In implementations, we typically express $A_B$ or $A_B^{-1}$ as a product of elementary
matrices.

In practice, good implementations of the Simplex Algorithm terminates after $2m$ to
$\sfrac{n}{2}$ iterations. Each iteration is very fast.

It is an open problem whether there exists a variant of Simplex Algorithm which is
guaranteed to terminate in at most $pn^q$ iterations for LP problems in SEF with $n$
variables, where $p$, $q$ are constants.

\makeheading{2019-10-10}

Midterm 1 was written on this day, as a result no classes were
held.
