\makeheading{2019-09-19}
\begin{defbox}
    \begin{definition}
        A \emph{non-linear program} has the form
        \[\min f(\bm{x})\]
        subject to
        \begin{align*}
            g_1(\bm{x}) & \leqslant 0 \\
            g_2(\bm{x}) & \leqslant 0 \\
                        & \vdots      \\
            g_m(\bm{x}) & \leqslant 0
        \end{align*}
        where
        $f:\mathbb{R}^n\rightarrow \mathbb{R}$,
        $g_i:\mathbb{R}^n\rightarrow\mathbb{R}$, $\forall i\in\{1,\dots,m\}$.
    \end{definition}
\end{defbox}

Every LP problem is a very special case of a NLP problem. IP problems can
also be formulated as NLP problems.
\begin{exbox}
    \begin{example}[Formulating LP Problems as NLP Problems]
        \begin{align*}
            x_i\in\mathbb{Z} & \iff \sin(\pi x_i)=0              \\
                             & \iff [\sin(\pi x_i)]^2\leqslant 0
        \end{align*}
        \begin{align*}
            x_i\in \{0,1\} & \iff x_i(1-x_i)=0              \\
                           & \iff x_i^2(1-x_i)^2\leqslant 0
        \end{align*}

        NLP problems have huge modelling power, as a result, one must understand the
        structure of the underlying problem and construct ``good'' NLP models that are
        amendable to analysis and solution techniques.
    \end{example}
\end{exbox}
\begin{exbox}
    \begin{example}[Portfolio Optimization]\label{Portfolio}
        There are $n$ stocks $1,\dots,n$ to invest in. We have a budget of $B$ dollars.
        We have an expected return (for \$1 investment at the end of our planning
        horizon) of $\mu_1,\dots,\mu_n$. We are also given
        $V\in \mathbb{R}^{n\times n}$, a variance covariance matrix so that if
        we invest in $x_1,\dots,x_n$ dollars in $n$ stocks, $1,\dots,n$ respectively,
        then the expected risk of such an investment is given by
        $\bm{x}^\top  V\bm{x}$.
        \[\sum\limits_{i = 1}^{n}\sum\limits_{j = 1}^{n}V_{ij}x_i x_j\]
        $x_j:=$ amount of investment in stock $j$ in dollars.

        Suppose we are also given a goal $G$ (a dollar amount we want as the value
        of our portfolio at the end of the planning horizon).

        \textbf{Data}
        \begin{itemize}
            \item Budget (\$) $\rightarrow$ B
            \item Goal (\$) $\rightarrow$ G
            \item Expected return $\rightarrow$ $ (\mu_1,\ldots,\mu_n)^\top $
            \item Variance-covariance matrix $\rightarrow$
                  $V\in \mathbb{R}^{n\times n}$
        \end{itemize}
        We want to minimize the risk of our portfolio while satisfying the budget and
        the goal constraints.

        (NLP)
        \[\min \bm{x}^\top  V\bm{x}\]
        subject to
        \begin{align*}
            \sum_{j = 1}^{n}x_j\leqslant B        \\
            \sum_{j = 1}^{n}\mu_j x_j\geqslant  G \\
            \bm{x}\geqslant  \bm{0}
        \end{align*}
        There are many variants of such models and extensions. For example, instead
        of a goal $G$, we may given an upper bound on the risk, say
        $R\in\mathbb{R}_{>0}$.

        (NLP)
        \[\max \sum_{j = 1}^{n}\mu_j x_j\]
        subject to
        \begin{align*}
            \sum_{j = 1}^{n}x_j\leqslant B  \\
            \bm{x}^\top  V\bm{x}\leqslant R \\
            \bm{x}\geqslant  \bm{0}
        \end{align*}
        We can handle many more variants and extensions. Suppose investing in stock
        $j$ below $\ell_j$ dollars is not allowed. For diversity of our portfolio, we
        want to invest in at least 20 stocks, and for the sake of simplicity we want
        to invest in at most 150 stocks. We introduce a binary-valued variable $z_j$.
        \[z_j:=
            \begin{cases}
                1 \text{, if we invest in stock $j$} \\
                0 \text{, otherwise}
            \end{cases}
            \forall j\in\{1,\dots,n\}\\
        \]
        Add these constraints:
        \begin{align*}
             & \ell_j z_j\leqslant x_j\leqslant B z_j\qquad \forall j\in\{1,\dots,n\}        \\
             & 20\leqslant \sum_{j = 1}^{n}z_j\leqslant 150 \qquad \forall j\in\{1,\dots,n\} \\
             & z_j\in\{0,1\} \qquad \forall j\in\{1,\dots,n\}                                \\
        \end{align*}
    \end{example}
\end{exbox}
