\makeheading{2020-02-05}
\section{Decoding Single-Error Correcting Codes}
Let $ H $ be a PCM for an $ (n,k) $-code $ C $ over $ GF(q) $
with $ d\geqslant 3 $.
\begin{align*}
    H\symbf{r}^\top
     & =H(\symbf{c}+\symbf{e})^\top                                                                          \\
     & =H\symbf{c}^\top+H \symbf{e}^\top                                                                     \\
     & =H\symbf{e}^\top                  & \quad & \text{since $ \symbf{c}^\top $ is in null space of $ H$ }
\end{align*}

\begin{Definition}{Syndrome}{syndrome}
    Let $ H $ be a parity-check matrix for an $ (n,k) $-code.
    The \textbf{syndrome} $ \symbf{s} $ of $ \symbf{r} $
    is defined to be $ \symbf{s}=H\symbf{r}^\top $.
\end{Definition}

\textbf{Notes}:
\begin{enumerate}[label=(\arabic*)]
    \item $ \symbf{r} $ and $ \symbf{e} $ have the same syndrome.
    \item If $ \symbf{e}=\symbf{0} $, then $ H\symbf{e}^\top=\symbf{0} $.
    \item If $ w(\symbf{e})=1 $, say $ \symbf{e}=(0,\ldots,0,\alpha,0,\ldots,0) $
          where $ \alpha $ is in the $ i^{\text{th}} $ position with $ \alpha \neq 0 $,
          then $ H\symbf{e}^\top=\alpha h_i $ where $ h_i $ is the $ i^{\text{th}} $
          column of $ H $.
    \item The converse of (2) and (3) are false.
\end{enumerate}

\begin{algorithm}
    \DontPrintSemicolon{}
    \caption{Decoding Algorithm for Single-Error Correcting Codes}\label{alg:Decoding Algorithm for Single-Error Correcting Codes}
    \SetKwInOut{Input}{Input}
    \SetKwInOut{Output}{Output}

    \Input{Parity-check matrix $ H = (\symbf{h}_1,\ldots ,\symbf{h}_n)^\top $
        and received vector $ \symbf{r} $}
    \Output{Decoded vector}
    $ \symbf{s}\gets H\symbf{r}^\top $\;
    \If{$ w(\symbf{s})=0 $} {
        \Return{$ \symbf{r} $}
    }
    \For{$ i\gets 0$ \KwTo{} $ n $} {
        \If{$ \symbf{s}=\alpha \symbf{h}_i $ with $ \alpha\neq 0 $} {
            \Return{$ \symbf{r}-\symbf{e} $}
        }
    }
    \Return{}
\end{algorithm}
\begin{Remark}{}{}
    $ \symbf{h}_i $ are column vectors in the input for each $ i\in[1,n] $.
\end{Remark}

\textbf{Claim}: If $ w(\symbf{e})\leqslant 1 $, then the decoding algorithm always
makes the correct decision.

\textbf{Note}: If $ H $ is a Hamming code and $ w(\symbf{e})\geqslant 2 $,
then this decoding algorithm will always make the wrong decision.

\begin{Example}{Single-Error Decoding}{}
    Consider the $ (7,4,3) $-binary Hamming code with PCM
    \[ H=\begin{bmatrix}
            1 & 0 & 1 & 0 & 1 & 0 & 1 \\
            0 & 1 & 1 & 0 & 0 & 1 & 1 \\
            0 & 0 & 0 & 1 & 1 & 1 & 1
        \end{bmatrix}_{3\times 7} \]
    Decode $ \symbf{r}=(0111110) $.

    \textbf{Solution.}
    \begin{enumerate}
        \item $ \symbf{s}=H\symbf{r}^\top=(011) $.
        \item $ \symbf{s} $ is the 6th column of $ H $, so $ \symbf{e}=(0000010) $.
        \item $ \symbf{c}=\symbf{r}-\symbf{e}=(0111100) $.
    \end{enumerate}
    Verify that $ H\symbf{c}^\top=\symbf{0} $.
\end{Example}

\textbf{General Decoding Problem for Binary Linear Codes}

\textbf{Instance}: An $ (n-k)\times n $ matrix $ H $ over $ GF(2) $
with $ \rank(H)=n-k $. $ \symbf{r}\in V_n(GF(2)) $.

\textbf{Find}: A vector $ \symbf{e}\in V_n(GF(2)) $ of minimum weight
with $ H\symbf{r}^\top=H\symbf{e}^\top $.

\textbf{Fact}: This problem is NP-hard.

\begin{itemize}
    \item P = problems solvable in polynomial time; that is, efficiently.
    \item NP = a certain class of problems including problems of strong
          practical interest which we do not know how to solve efficiently.
    \item NP-hard = If any single problem in this class of problems can be
          solved efficiently, then so can all problems in NP, in which case P=NP\@.
\end{itemize}
