\makeheading{ 2020-01-10 }

\section{Decoding Strategy}
Suppose we have an $ [n,M] $-code $ C $ over $ A $ of distance $ d $.
We need to adopt a strategy for the channel decoder (henceforth called the decoder).
When the decoder receives an $ n $-tuple $ \bm{r}\in A^n $ it must make some decision.
This decision may be one of
\begin{enumerate}[label=(\roman*)]
    \item no errors have occurred; accept $ \bm{r} $ as a codeword.
    \item errors have occurred; correct $ \bm{r} $ to a codeword $ \bm{c} $;
          e.g. $ 0 \rightarrow 0000,\,1 \rightarrow 1111,\,r=0001 $ corrected to
          $ 0000 $.
    \item errors have occurred; no correction is possible.
\end{enumerate}

\section{Nearest Neighbour Decoding}
\subsection{Incomplete Maximum Likelihood Decoding (IMLD)}
Correct $ \bm{r} $ to the unique codeword $ \bm{c} $ for which
$ d(\bm{r},\bm{c}) $ is smallest.
If $ \bm{c} $ is not unique, reject $ \bm{r} $.
\subsection{Complete Maximum Likelihood Decoding (CMLD)}
Same as IMLD, except ties are broken arbitrarily.

\textbf{Question:} Is IMLD a reasonable strategy?

\begin{Theorem}{}{}
    IMLD selects the codeword $ \bm{c} $ that maximizes $ P(\bm{r}\mid \bm{c})$; that is,
    it maximizes the probability $ \bm{r} $ is received, given $ \bm{c} $ was sent.
\end{Theorem}

We actually want to maximize $ P(\bm{c}\mid \bm{r}) $, but we will ignore that for now.

\begin{proof}
    Suppose $ \bm{c}_1,\bm{c}_2\in C $ with $ d(\bm{c}_1,\bm{r})=d_1 $
    and $ d(\bm{c}_2,\bm{r})=d_2 $.
    Suppose $ d_1>d_2 $. Now,

    $ P(\bm{r}\mid \bm{c}_1)=(1-p)^{n-d_1}\left( \frac{p}{q-1} \right)^{d_1} $ and
    $ P(\bm{r}\mid \bm{c}_2)=(1-p)^{n-d_2}\left( \frac{p}{q-1} \right)^{d_2} $.

    Hence,
    \begin{align*}
        \frac{P(\bm{r}\mid \bm{c}_1)}{P(\bm{r}\mid \bm{c}_2)}
         & =(1-p)^{d_2-d_1}\left( \frac{p}{q-1}  \right)^{d_1-d_2} \\
         & =\left[ \frac{p}{(1-p)(q-1)}  \right]^{d_1-d_2}
    \end{align*}
    Recall that, for a $ q $-ary channel, we can assume that $ p<\frac{q-1}{q} $. Thus,
    \begin{align*}
         & \implies pq < q-1                 \\
         & \implies 0 < q-1-pq               \\
         & \implies p < q-1-pq+p             \\
         & \implies p < (1-p)(q-1)           \\
         & \implies \frac{p}{(1-p)(q-1)} < 1
    \end{align*}
    Since $ d_1>d_2 $, we get $ \frac{P(\bm{r}\mid \bm{c}_1)}{P(\bm{r}\mid \bm{c}_2)} < 1 $, and so
    $ P(\bm{r}\mid \bm{c}_1)<P(\bm{r}\mid \bm{c}_2) $.
\end{proof}
The ideal strategy is to correct $ \bm{r} $ to $ \bm{c}\in C $ such that
$ P(\bm{c}\mid \bm{r}) $ is maximized. This is \textbf{Minimum Error Decoding (MED)}.


\begin{Example}{IMLD $ \neq $ MED}{}
    Let $ C=\{\underbrace{000}_{\bm{c}_1},\underbrace{111}_{\bm{c}_2}\},\,
        P(\bm{c}_1)=0.1,\, P(\bm{c}_2)=0.9,\, p=\sfrac{1}{4} $, and $ \bm{r}=100 $.

    \textbf{IMLD} $ \bm{r} $ is decoded to to $ \bm{c}_1 = 000 $.

    \textbf{MED}
    \begin{align*}
        P(\bm{c}_1\mid \bm{r})
         & =\frac{P(\bm{r}\mid \bm{c}_1)P(\bm{c}_1)}{P(\bm{r})} \\
         & =\frac{p(1-p)^2(0.1)}{P(\bm{r})}                     \\
         & =\frac{0.0140625}{P(\bm{r})}
    \end{align*}
    \begin{align*}
        P(\bm{c}_2\mid \bm{r})
         & =\frac{P(\bm{r}\mid \bm{c}_2)P(\bm{c}_2)}{P(\bm{r})} \\
         & =\frac{p^2(1-p)(0.9)}{P(\bm{r})}                     \\
         & =\frac{0.0421875}{P(\bm{r})}
    \end{align*}
    Since $ P(\bm{c}_1\mid \bm{r})<P(\bm{c}_2\mid \bm{r}) $,
    $ \bm{r} $ is decoded to $ \bm{c}_2=111 $.
\end{Example}


\textbf{Notes:}
\begin{enumerate}[label=(\roman*)]
    \item IMLD selects $ \bm{c} $ such that $ P(\bm{r}\mid \bm{c}) $ is maximum.
    \item MED selects $ \bm{c} $ such that $ P(\bm{c}\mid \bm{r}) $ is maximum.
    \item MED has a drawback that it requires knowledge of $ P(\bm{c}_i) $ for
          each $ i\in [1,M] $.
    \item Suppose source messages are equally likely, so
          $ P(\bm{c}_i)=\frac{1}{M} $ for each $ i\in[1,M] $.
          Then,
          \[ P(\bm{r}\mid \bm{c}_i)=\frac{P(\bm{c}_i\mid \bm{r})P(\bm{r})}{P(\bm{c}_i)}
              =P(\bm{c}_i\mid \bm{r}) \underbrace{M P(r)}_{\text{constant}}\]
          So, maximizing $ P(\bm{r}\mid \bm{c}_i) $ is the same as maximizing
          $ P(\bm{c}_i\mid \bm{r}) $. Thus, IMLD is the same as MED in this case.
\end{enumerate}
In the remainder of the course, we will use IMLD/CMLD\@.
