\makeheading{ 2020-01-17 }

\begin{Definition}{Isomorphic}{isomorphic}
    We say two fields $F$ and $S$ are \textbf{isomorphic} if they have the same
    binary operations and if there exists a bijection between them.
\end{Definition}

\begin{Definition}{Subfield}{subfield}
    Let $ F $ be a field. A subset $ S\subseteq F $ is called a \textbf{subfield}
    of $ F $ if $ S $ is a field itself with respect to the same operations
    of $ F $.
\end{Definition}

\begin{Example}{Subfield}{}
    Let $ F $ be a finite field where $ \ch(F)=p $. Consider
    $ E=\set{0,1,1+1,\ldots,\underbrace{1+\cdots+1}_{p-1}}\subseteq F $.
    We see that $ E $ is a field with the same field operations as $ F $.
    Also, $ E $ has order $ p $. If we label the elements of $ E $
    naturally such that $ \underbrace{1+\cdots+1}_{p-1}
        \longleftrightarrow p-1 $,
    then
    \[ E=\set{0,1,1+1,\ldots,\underbrace{1+\cdots+1}_{p-1}}
        = \mathbb{Z}_p= \set{0,1,2,\ldots ,p-1}\subseteq F \]
    So $ E $ is isomorphic to $ \mathbb{Z}_p $.
\end{Example}

\begin{Theorem}{}{subfield_thm}
    If $ F $ is a finite field of characteristic $ p $, then
    $ \mathbb{Z}_p $ is a subfield of $ F $.
\end{Theorem}

\begin{Exercise}{}{}
    Prove~\Cref{thm:subfield_thm}.
\end{Exercise}

\begin{Definition}{}{}
    Let $ F $ be a finite field, and consider $ \mathbb{Z}_p\subseteq F $.
    \begin{itemize}
        \item Each $ v\in F $ is vector.
        \item Each $ c\in\mathbb{Z}_p $ is a scalar.
        \item Addition in $ F $ is defined by vector addition.
        \item Multiplication in $ F $ by elements in $ \mathbb{Z}_p $
              is defined by scalar multiplication.
    \end{itemize}
\end{Definition}

\begin{Theorem}{}{field_vec_space}
    If $ F $ is a finite field of characteristic $ p $, then $ F $
    is a vector space over $ \mathbb{Z}_p $.
\end{Theorem}

\begin{Exercise}{}{}
    Prove~\Cref{thm:field_vec_space}.
\end{Exercise}

\begin{Theorem}{}{ord_pk}
    If $ F $ is a finite field of characteristic $ p $, then
    \[ \ord(F)=p^k \]
    for some $ k\in\mathbb{Z}_{\geqslant 1} $.
\end{Theorem}

\begin{Proof}{\Cref{thm:ord_pk}}{}
    Let $ k $ be the dimension of the vector space $ F $ over $ \mathbb{Z}_p $.
    Let $ \set{\alpha_1,\ldots ,\alpha_k} $ be a basis for $ F $. Then, every element
    in $ F $ can be written as
    \[ c_1\alpha_1+\cdots+c_k\alpha_k \]
    where $ c_i\in\mathbb{Z}_p $. For each $ \alpha_i $, there are $ p $
    possible choices for $ c_i $, hence $ \ord(F)=p^k $.
\end{Proof}

\begin{Example}{}{}
    There is no field of order $ 6 $.
\end{Example}

\textbf{Question}: Is there a finite field of order $ 4,\,8,\,9 $?

\section{Irreducible Polynomials}

\begin{Definition}{Set of all polynomials in $ x $ over $ F $}{}
    Let $ F $ be a field. The \textbf{set of all polynomials in $ x $ over $ F $}
    (polynomials with coefficients from $ F $) is denoted $ F[x] $. Addition
    and multiplication are both done in the usual way, with coefficient arithmetic
    in $ F $.
\end{Definition}

\begin{Example}{Polynomial Modular Arithmetic}{}
    In $ \mathbb{Z}_{11} $, $ (2+5x+6x^2)+(3+9x+5x^2)=5+3x $.
\end{Example}

\begin{Theorem}{}{}
    Let $ F $ be a field. $ F[x] $ is a commutative ring.
\end{Theorem}

\begin{Definition}{}{}
    Let $ F $ be a field and let $ f\in F[x] $ with $ \deg(f)\geqslant 1 $.
    If $ g,h\in F[x] $ with $ f\mid (g-h) $, then we write
    \[ g\equiv h \pmod{f} \]
    or equivalently, we can write $ g-h=\ell f $ for some $ \ell\in F[x] $.
\end{Definition}

\begin{Theorem}{}{}
    Congruence is an equivalence relation.
\end{Theorem}

\begin{Definition}{Equivalence class containing $\symbf{g\in F[x]}$}{}
    For a given $ f\in F[x] $, the \textbf{equivalence class containing $\symbf{g\in F[x]}$}
    is
    \[ [g]=\set{h\in F[x]: h\equiv g\pmod{f}} \]
\end{Definition}

\begin{Definition}{}{}
    For $ g,h\in F[x] $, we define addition and multiplication as follows:
    \begin{itemize}
        \item Addition: $ [g]+[h]=[g+h] $
        \item Multiplication: $ [g][h]=[gh] $
    \end{itemize}
\end{Definition}

\begin{Theorem}{}{equiv_class}
    \begin{enumerate}[label=(\arabic*)]
        \item The set of all equivalence classes, denoted $ F[x]/(f) $
              where $ f\in F[x] $ and $ \deg(f)\geqslant 1 $ is a
              commutative ring.
        \item The polynomials in $ F[x] $ of degree less than degree of $ f $
              are a system of distinct representatives of equivalence classes in
              $ F[x]/(f)$.
    \end{enumerate}
\end{Theorem}

\begin{Proof}{\Cref{thm:equiv_class}}{}
    Let $ g\in F[x] $. By division algorithm for polynomials we can write
    $ g=\ell f+r $ where $ \deg(r)<\deg(f) $. So, $ g-r=\ell f $. Hence,
    $ g\equiv r\pmod{f} $. Thus, $ [g]=[r] $ and we have $ \deg(r)<\deg(f) $.
    Also, if $ r_1,r_2\in F[x] $ with $ r_1\neq r_2 $, and
    $ \deg(r_1),\deg(r_2)<\deg(f) $, then
    \[ f\nmid (r_1-r_2)\iff r_1\not\equiv r_2\pmod{f} \]
    Thus, $ [r_1]\neq [r_2] $.
\end{Proof}
