\input{../../../../preamble.tex}

\begin{document}

4. (10 marks) \textbf{Telephone numbers}\\
Let $C$ be an $[n,M]$-code with distance $d$ over an alphabet $A$
of size $q$. Let $e = \lfloor \frac{d-1}{2} \rfloor$.

\textbf{(a) Prove that $M \sum_{i=0}^e {n \choose i} (q-1)^i \leq q^n$.\\
     (Hint: Consider the number of words in a sphere of radius $e$ centred
      at a codeword)}
\begin{proof}
    Consider the number of words in a single sphere of radius $ e $ centred
    at a codeword $ c\in C $, call it $ x $. Every pair of spheres do not intersect
    since $ d(C)=d\implies C $ is a $ e $-error correcting code. There are up to
    $ \binom{n}{e} $ (from $ \binom{n}{0} $) ways
    to choose the components of $ c $ to differ by $ (q-1) $ other values.
    Hence,
    \[ x=\sum\limits_{i=0}^{e} \binom{n}{i}(q-1)^i \]
    For the total number of spheres, $ M $, we get
    \[ M x = M \sum\limits_{i=0}^{e} \binom{n}{i}(q-1)^i \]
    which cannot possibly be more than the total number of words in the alphabet
    $ A $ of size $ q $, $ q^n $. Thus,
    \[ M \sum\limits_{i=0}^e {n \choose i} (q-1)^i \leq q^n \]
\end{proof}

\textbf{(b) Suppose that there are 110 million telephones in a country.
      Is it possible to assign 10-digit decimal numbers to these
      telephones so that a single error in dialing can always be
      (automatically) corrected? (Explain)}

      Try the following parameters with the sphere packing bound:
      \begin{itemize}
        \item $ M=110000000 $ (number of telephones/codewords)
        \item $ q=10 $ (all decimal digits)
        \item $ n=10 $ (length of each telephone number/length of words)
        \item $ e=1 $ (errors to be corrected)
      \end{itemize}
      We get,
      \[ (110000000)\sum\limits_{i=0}^1 {10 \choose i} (10-1)^i
        =10010000000
        \nleq 10^{10} \]

        The block code \textbf{does not} satisfy the \emph{necessary}
        sphere packing bound condition, thus it is \textbf{not possible} to assign a
        10-digit decimal numbers to these telephones so that a single error
        in dialing can always be (automatically) corrected.

\textbf{(c) Suppose that there are 80 million telephones in a country.
      Is it possible to assign 10-digit decimal numbers to these
      telephones so that a single error in dialing can always be
      (automatically) corrected? (Explain)}

      Try the following parameters with the sphere packing bound:
      \begin{itemize}
        \item $ M=80000000 $ (number of telephones/codewords)
        \item $ q=10 $ (all decimal digits)
        \item $ n=10 $ (length of each telephone number/length of words)
        \item $ e=1 $ (errors to be corrected)
      \end{itemize}
      We get,
      \[ (80000000)\sum\limits_{i=0}^1 {10 \choose i} (10-1)^i=
        7280000000
        \leq 10^{10} \]

        The block code \textbf{does} satisfy the \emph{necessary}
        sphere packing bound condition, but the bound is not \emph{sufficient}
        to say whether there exists a code for this problem. Thus,
        the solution to this problem is \textbf{inconclusive}.

\end{document}