\section{2020-03-06}
\underline{Roadmap}:
\begin{enumerate}[label=(\roman*)]
    \item Recap (excluded from these notes)
    \item Testing of hypotheses (Null vs Alternate) and (Two-sided vs One-sided tests)
    \item Clicker
\end{enumerate}
\underline{Hypothesis Testing}

\begin{Definition}{}{}
    A hypothesis is a statement about the (parameters of) population. There
    are two (competing) hypotheses.

    \underline{Null Hypothesis} $ H_0 $: current belief, conventional wisdom

    \underline{Alternate Hypothesis} $ H_1 $: challenger to the conventional wisdom
\end{Definition}



\begin{Example}{}{}
    Suppose we want to test whether a coin is biased. We flip the coin $ 100 $
    times and get $ 52 $ heads. Let $ \theta=P(H) $
    \begin{itemize}
        \item $ H_0 $: $ \theta=\frac{1}{2} $
        \item $ H_1 $: $ \theta\neq \frac{1}{2} $
    \end{itemize}
    Approach $ p $-value approach.
\end{Example}



\begin{Definition}{}{}
    The $ p $-value: is the probability of observing my evidence (or worse)
    under the assumption that $ H_0 $ is true. The lower the $ p $-value,
    the strong is the evidence against $ H_0 $.
\end{Definition}

\underline{Notes}:
\begin{itemize}
    \item $ H_0 $ and $ H_1 $ are not treated symmetrically.
    \item Unless there is overwhelming evidence (``beyond a reasonable doubt'')
          against $ H-0 $, we stick with it. The burden is on the challenger.
          \begin{center}
              \begin{tabular}{| *{3}{>{\centering\arraybackslash}p{3cm} |}}
                  \hline
                                           & $ H_0 $ is true & $ H_1 $ is true \\ \hline
                  Reject $ H_0 $ (convict) & $ X_1 $         & $ \checkmark $  \\ \hline
                  Do not reject $ H_0 $    & $ \checkmark $  & $ X_2 $         \\ \hline
              \end{tabular}
          \end{center}
          where $ X_1 $ is a Type I error and $ X_2 $ is a Type II error.
\end{itemize}

\underline{Two-sided vs One-sided tests}:
\begin{itemize}
    \item $ H_0 $: $ \theta=\frac{1}{6} $
    \item $ H_1 $: $ \theta<\frac{1}{6} $
\end{itemize}

\underline{Clicker Question} The $ p $-value $ = P(H_0\text{ is true}) $.
\begin{enumerate}[label=(\alph*)]
    \item True
    \item \textbf{False}
\end{enumerate}
