\section{2020-02-10}
\underline{Roadmap}:
\begin{itemize}
    \item Interval Estimation
          \subitem Likelihood Estimation
          \subitem Confidence Intervals: Coverage probabilities, Pivotal Quantities
\end{itemize}
\begin{exbox}
    \begin{example}
        The approval rating of Trump is $ 49\% $ ($ 49\% $ is the most ``likely'' value of $ \theta $)
        where $ \theta= $ population approval rating.
        \begin{itemize}
            \item What is the ``Margin of Error''?
                  \subitem How does one calculate it?
        \end{itemize}
    \end{example}
\end{exbox}
\underline{Setup} $ Y_1,\ldots ,Y_n $ are iid random variables with
distribution (density) $ f(y;\theta) $ where $ \theta= $ unknown attribute.

\underline{Objective}: Based on our data $ \{y_1,\ldots ,y_n\} $, we would
construct an interval $ [a,b] $
\[ a(y_1,\ldots ,y_n),b(y_1,\ldots ,y_n) \]
which are the ``reasonable'' values of $ \theta $.

\underline{Method 1}: Through the relative likelihood function.

\underline{Intuition}: $ \theta $ is ``reasonable'' of $ L(\theta) $
is ``close'' to $ L(\hat{\theta}) $, where $ \theta= $ MLE.

\begin{defbox}
    \begin{definition}
        A $ 100p\% $ likelihood interval for $ \theta $ where $ p\in[0,1] $
        \[ \{\theta:R(\theta)\geqslant p\} \]
    \end{definition}
\end{defbox}
Take $ p=0.5 $, we get that $ R(\theta)\geqslant 0.5 $, so
\[ \implies L(\theta)\geqslant 0.5 L(\hat{\theta}) \]
The value of the likelihood at $ \theta $ is at least $ 50\% $ of the value of the
likelihood evaluated at the MLE.

\underline{Convention}
\begin{itemize}
    \item $ R(\theta)\geqslant 0.5\implies \theta $ is very plausible
    \item $ 0.1\leqslant R(\theta)<0.5\implies \theta $ is plausible
    \item $ 0.01\leqslant R(\theta)<0.1\implies \theta $ is implausible
    \item $ R(\theta)<0.01\implies \theta $ is very implausible
\end{itemize}
\begin{exbox}
    \begin{example}
        A coin is tossed $ 200 $ times and we observe $ 120 $ heads.
        Let $ \theta=P(\text{H}) $. Is $ \theta=0.5 $ plausible?

        \textbf{Solution.} Find the $ 10\% $ likelihood interval for $ \theta $.
        \[ L(\theta)=\binom{200}{120}\theta^{120}(1-\theta)^{80} \]
        We are given that $ \hat{\theta}=0.6 $.
        \[ \left\{ \theta:\frac{\theta^{120}(1-\theta)^80}{0.6^{120}(0.4)^{80}}\geqslant 0.1 \right\} \]
        Thus,
        \[ R(\theta)=\frac{\theta^{120}(1-\theta)^80}{0.6^{120}(0.4)^{80}} \]
        Is $ \theta=0.5 $ plausible? Plug in $ \theta=0.5 $ and check if $ R(0.5)\geqslant 0.1 $.
    \end{example}
\end{exbox}
\begin{exbox}
    \begin{example}
        Two Binomial experiments.
        \begin{itemize}
            \item $ n_1=1000 $, $ y_1=200 $
            \item $ n_2=100 $, $ y_2=20 $
            \item $ y= $ number of successes
            \item $ n= $ number of trials
        \end{itemize}
        Which $ 10\% $ likelihood interval is wider?

        \textbf{Solution.} We have $ \hat{\theta}=0.2 $.
        $ n=100 $ yields a wider interval.
    \end{example}
\end{exbox}

\underline{Method 2}: Confidence intervals.

\underline{Setup}: There is a pre-specified probability (coverage probability),
say $ 95\% $ or $ 99\% $ for example.

\underline{Objective}: Based on your data, we want to estimate the (random)
interval which would contain $ \theta $ with that probability.
\begin{exbox}
    \begin{example}
        The STAT 231 scores of UW Math students is normally distributed independently
        \[ Y_i \thicksim N(\mu,64) \]
        A sample of $ 25 $ students are collected
        \[ \overline{y}=75 \]
        Find the $ 95\% $ confidence interval for $ \mu $.
    \end{example}
\end{exbox}
\underline{Sampling Distributions}

\underline{Idea}: All the data summaries are also outcomes of some random experiment.
\[ Y_1,\ldots ,Y_n \thicksim N(\mu,\sigma^2) \quad\text{iid}\]
\[ \implies \overline{Y} \thicksim N(\mu,\sigma^2/n) \]
Our sample mean $ \overline{y} $ is an outcome of this experiment.
