---
output:
  pdf_document: default
  header-includes: \usepackage{amsmath}
editor_options: 
  chunk_output_type: console
---

```{r}
## NASA rocket data example

## From: R.S. Jankovsky, T.D. Smith, A.J. Pavli (1999). "High-Area-Ratio Rocket
## Nozzle at High Combustion Chamber Pressure-Experimental and Analytical
## Validation".

# setwd(...) first if your CSV file is somewhere else
rocket <- read.csv("csv/rocket.csv")
# output all data in rocket vector
rocket
```

$Y$ (thrust) is the response variable, and there
are two explanatory variables $x_1,x_2$
(nozzle, propratio) where nozzle is coded
as 1 if it's large.

```{r}
# Scatter plots where mfrow is used to put multiple plots on one image
par(mfrow = c(1, 2))
plot(rocket$nozzle,
     rocket$thrust,
     ylab = "Thrust",
     xlab = "Nozzle size (1 = large)")
plot(rocket$propratio,
     rocket$thrust,
     ylab = "Thrust",
     xlab = "Propellant to fuel ratio")
```

Left is nozzle size vs thrust. Right is propellant
relationship vs thrust.

```{r}
# Fit MLR using lm
m1 <- lm(thrust ~ nozzle + propratio, data = rocket)
summary(m1)
m2 <- lm(thrust ~ 0 + nozzle, data = rocket)
summary(m2)
anova(m1)
```

On the left it's $Y$ (response variable) and on the
right it's $x_1,x_2$ (explanatory variables).
From summary, we get the estimate vector
$\hat{\beta}=(473.6039, 16.7383,-1.0948)^\top$.

```{r}
# Manual beta estimates where rep is used to make the columns of 1s
X <- cbind(rep(1, 12), rocket$nozzle, rocket$propratio) # X matrix
y <- matrix(rocket$thrust, ncol = 1) # response vector
beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y
beta_hat
```

`solve` is used for the inverse. `%*%` is used
for matrix-matrix multiplication, and `t(X)`
is used for transposing $X$.

```{r}
# Manual sigma estimate
mu_hat <- X %*% beta_hat # fitted values
e <- y - mu_hat # residuals
sigma_hat <- sqrt((t(e) %*% e) / 9) # Note n-p-1 = 12-2-1 = 9
sigma_hat
sigma_hat <- sqrt(sum(e ^ 2) / 9) # equivalent
sigma_hat
```
\begin{itemize}
  \item $\hat{\boldsymbol{\mu}}=X\hat{\boldsymbol{\beta}}$
  \item $\boldsymbol{e}=\boldsymbol{y}-\hat{\boldsymbol{\mu}}$
  \item $\hat{\sigma}=\sqrt{\biggl(\sum_{i=1}^{n}e_i^2\biggr)/9}=2.6545$, or
  \item $\hat{\sigma}=\sqrt{(\boldsymbol{e}^\top\boldsymbol{e})/9}=2.6545$
\end{itemize}
```{r}
# Covariance matrix of beta_hat
vcov(m1)
sqrt(diag(vcov(m1))) # SEs of individual betas

# Manual
se_beta <- sigma_hat * sqrt(diag(solve(t(X) %*% X)))
se_beta
```

\begin{itemize}
  \item $Se(\hat{\boldsymbol{\beta}})=\hat{\sigma}\sqrt{(X^\top X)^{-1}}=(4.71, 1.53, 0.94)^{\top}$
\end{itemize}

```{r}
# Estimate the mean response for units with small nozzle and propellant ratio 5.5
# include a 95% CI
predict(
  object = m1,
  newdata = data.frame(nozzle = 0, propratio = 5.5),
  interval = "confidence",
  level = 0.95
)
```

Therefore, $\hat{y}_0=467.58$. The 95\% confidence interval for the mean
response given $\boldsymbol{x}_0$ is $[464.7929,470.3719]$.

```{r}
# Manual calculation
x0 <- matrix(c(1, 0, 5.5), nrow = 1)
y0_hat <- x0 %*% beta_hat
y0_hat
# mu0 is also known as \hat{Y}_0
se_mu0 <- sigma_hat * sqrt(x0 %*% solve(t(X) %*% X) %*% t(x0))
se_mu0
crit_val <- qt(0.975, 9)
ci_lo <- y0_hat - crit_val * se_mu0
ci_hi <- y0_hat + crit_val * se_mu0
c(y0_hat, ci_lo, ci_hi)
```

\begin{itemize}
  \item $\boldsymbol{x}_0=\begin{bmatrix}1&0&5.5\end{bmatrix}$
  \item $\hat{y}_0=\boldsymbol{x}_0\hat{\boldsymbol{\beta}}=467.5824$
  \item $Se(\hat{Y}_0)=\hat{\sigma}\sqrt{\boldsymbol{x}_0(X^\top X)^{-1}\boldsymbol{x}_0^\top}=
  1.233132$
\end{itemize}
Therefore, $\hat{y}_0=467.58$. The 95\% confidence interval for the mean
response given $\boldsymbol{x}_0$ is $[464.7929,470.3719]$.

```{r}
# Predict the value of the response for a unit with small nozzle and propellant ratio 5.5
# include a 95% PI
predict(
  object = m1,
  newdata = data.frame(nozzle = 0, propratio = 5.5),
  interval = "prediction",
  level = 0.95
)
```

Therefore, $y_0=467.5824$. The 95\% prediction interval for the
response $(y_0)$ given $\boldsymbol{x}_0$ is $[460.9612 474.2036]$.

```{r}
# Manual calculation for an individual
x0 <- matrix(c(1, 0, 5.5), nrow = 1)
y0_hat <- x0 %*% beta_hat
se_y0 <- sigma_hat * sqrt(1 + x0 %*% solve(t(X) %*% X) %*% t(x0))
se_y0
crit_val <- qt(0.975, 9)
pi_lo <- y0_hat - crit_val * se_y0
pi_hi <- y0_hat + crit_val * se_y0
c(y0_hat, pi_lo, pi_hi)
```

\begin{itemize}
  \item $Se(Y_0-\hat{Y}_0)=\hat{\sigma}\sqrt{1+\boldsymbol{x}_0(X^\top X)^{-1}\boldsymbol{x}_0^\top}=2.926941$
\end{itemize}
